Data Analysis Project

### 2023-03-18 products/manuscript/Manuscript.qmd
update mansucript

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
#library(here)
#library(knitr)
```

# Summary/Abstract
_Write a summary of your project._

@@

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._

@@

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]

@@

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

## Statistical analysis
_Explain anything related to your statistical analyses._

@@

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

@@

@@

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@@

@@

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

@@

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

>>>>

---
title: "Data Analysis Project"
subtitle: ""
author: Leah Lariscy
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract

*Write a summary of your project.*

@@

## General Background Information

Infection by Severe Acute Respiratory Syndrome Coronavirus 2 (SARS-CoV-2), causing COVID-19, is followed by the shedding of viral particles by multiple excretory functions, including stool and urine production.

Subsequently, these viral particles can be detected in wastewater influent via RNA extraction, followed by Reverse Transcriptase Quantitative Polymerase Chain Reaction (RT-qPCR) utilizing SARS-CoV-2 specific primers.

Wastewater-based detection methods have been utilized across the globe as an independent indicator of SARS-CoV-2 viral prevalence and, depending on community-specific factors, as a leading indicator of clinical case trends \[4\]. WBE methods may also become a primary tool in understanding transmission dynamics as case under reporting increases.

## Description of data and data source

24-hour composite samples collected twice weekly, stored at 4C until ready for extraction. 6 extraction replicates were performed for each collection date at treatment facility.

2019-nCoV N1 and N2 primer-probe sets were used for universal detection of SARS-CoV-2.

The concentration per reaction of each gene target was determined by the corresponding standard curve, then sample concentration in copies per liter was determined by Equation 1. Total daily copies per day per WRF was determined by Equation 2.

## Questions/Hypotheses to be addressed

Is there a relationship between viral load and clinical case observations?

Is there a relationship between qPCR percent positive and clinical case observations?

What do viral load trends look like over time?

What do clinical case trends look like over time?

How do each of these temporal trends compare to one another over time?

@@

@@

## Data aquisition

Wastewater SARS-CoV-2 data was obtained through weekly surveillance efforts utilizing RT-qPCR methods.

Water Reclamation Facility data was obtained through data share directly from the WRF representatives.

COVID data downloaded from Georgia DPH website.

## Data import and cleaning

*qPCR data*

WWTP data

*DPH COVID data*

## Statistical analysis

Simple analyses: Wastewater and clinical data over time, linear regression model of wastewater vs clinical case data

Full analyses: Wastewater data as predictor for clinical case outcomes

Steps I took to build and tune model

@@

@@

![Distribution of recorded seven-day moving average for each day](images/7DMA_histo.png)

![Distribution of viral load per sample collection per target](images/total_copies_histo.png)

@@

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

@@

![Clinical case seven-day moving average over time](images/7DMA_time_plot.png)

![Wastewater viral load over time](images/avg_total_copies_time.png)

![Linear regression of viral load vs case seven-day moving average](images/total_copies_avg_vs_7DMA.png)

@@

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

@@

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
